Computer Science, Vignettes, and Citation Analysis

### 2019-12-16 Scientometrics/comp.tex
lots of changes through results, nothing changed in discussion

Computer science, and its applications, has experienced rapid growth and diversification over the last twenty years. As observed in a 2017 US National Academies Report, ``A wide range of jobs in virtually all sectors demand computing skills to an unprecedented extent. And every academic discipline finds itself incorporating computing into its research and educational mission"~\cite{nas_2017}. More recently, the collective influence of the Internet of Things (IoT), `big' data, accessible cloud computing, and advances in artificial intelligence have been postulated as a recent driver for growth and evolution~\cite{siebel2019_digital}. Given this powerful influence some understanding of the present state and structure of computer science and its relationship to other fields could inform planning and policy making at multiple levels from national level funding all the way down to faculty hiring strategy. For the same reason of broad influence, computer science can also serve as a model system for studying interdisciplinarity.

Noting that that the scientific literature serves a rich source of information to study the structure and historical development of a field, Salton and Bergmark conducted a historical study  of the computer science literature (419 computer articles published in 1974,  and 3,812 references cited in these articles)~\cite{salton_citation_1979}. These authors described the global structure of computer science as comprising three main areas (i) theoretical foundations, such as theory of computation, (ii) hardware and computer systems, such as architecture,  and (iii) software, such as programming systems.  Related areas noted were (a) mathematics of computing, such as numerical analysis, (b) special software topics, such as operating systems, (c) data management and database systems, (d) methodologies valid for multiple applications, such as algebraic manipulation, (e) computer applications, such as computer graphics, (f) non-technical aspects, such as computer education.

Looking beyond Salton and Bergmark's historical triad of theoretical foundations; hardware and computer systems; and software, the Computing Classification System (CCS) published by the Association for Computing Machinery now consists of 13 top-level areas that reflect a more current view of the field~\cite{acm_ref}. This classification also addresses relationships with other fields under the category of Applied Computing. However, an easy way to map scientific publications to the CCS, especially interdisciplinary articles or those from proximal fields does not seem to be available.

Other classification systems are available such as the All Science Journal Classification (ASJC) developed and maintained by Scopus, the Web of Science research and categorical classifications from Clarivate Analytics, and the NSF classification system~\cite{nsf_classification} [add citations for Scopus and Web of Science]. Scopus and the Web of Science also include comprehensive bibliographies with citation links and proprietary unique identifiers for publications. All three systems rely on applying one or more journal-derived labels to articles.  A logical prediction, given the diversity of articles within journals, is limited specificity at the article level. Shu and colleagues recently noted in a comparative study of the Chinese Science Citation Database (CSCD) and the Web of Science, that 46\% of articles did not belong to the discipline of the journal they were published in~\cite{shu_comparing_2019}. Others have also discussed and critiqued disciplinary assignments using journal based classifications~\cite{wang_large-scale_2016,perianes-rodriguez_comparison_2017}. Article classification systems have been constructed, which escape some of the criticisms of journal-based classification~\cite{traag_louvain_2019,boyack_classification_2014,waltman_new_2012} but seem to be in the research domain rather than the domain of practical use [need to check on this].[May need another bridging sentence to the next paragraph].

@@

@@

\subsection{Clustering methods} In applying both clustering by direct citation and by co-citation, we attempted to consider, wherever possible, the criteria articulated by \v{S}ubelj, van Eck, and Waltman~\cite{subelj_clustering_2016}, of (i) Cluster sizes- the largest cluster should be no more than 10 times the smallest one (ii) Small clusters- should be eliminated (iii) Clustering stability- small changes and replicates should yield similar results (iv) Computing time- minimized where possible (v) Intuitive sensibility- determined by qualitative examination.

\paragraph{Clustering by Direct Citation.} Graclus is a spectral graph clustering package that computes normalized cut and ratio association for a given undirected graph~\cite{graclus_2007}.  We used v1.2 in our experiments. The \emph{comp} dataset was formatted as an undirected graph, stored in a file with a header line indicating the number of nodes and edges, and used as input to Graclus, which requires the number of clusters to be formed as an input parameter. In preliminary experiments, we varied the number of clusters to be formed between 10 and 50 clusters (data not shown).  At around 20 clusters, clusters size was relatively stable with the largest cluster containing roughly 10 times the number of nodes in the smallest one~\cite{subelj_clustering_2016}. Consequently, we generated 20 clusters, labeled 0-19, using Graclus (Table~\ref{tab:graclus}).

We also used conductance, as defined in Shun et al.~\cite{shun_parallel_2016}, to  evaluate clustering (smaller is better), noting that conductance has been found to be ``the stand-alone quality metric that best indicates performance on information recovery metrics" \cite{emmons2016analysis}. Data from \emph{comp} were clustered into 18, 20, or 22 clusters and in each case the highest numbered cluster (17, 19, or 21) had the greatest conductance value and the smallest number of nodes, suggesting that it effectively served as a container for `left over publications' during the Graclus process. Thus, we limited our consideration of cluster 19 (the 20th cluster), also the smallest cluster, when interpreting results.  The remaining 19 clusters had conductance values ranging from 0.09 to 0.25 with a median conductance of 0.15, Figure~\ref{fig:graclus_comparison_conductance}, Table~\ref{tab:comp}.

\paragraph{Clustering by Co-Citation.}
For an alternate view of these DBLP data we constructed clusters using co-citation, the frequency with which a pair of articles is cited by other articles. To cluster by co-citation, we used a modification of variable level clustering combined with agglomerative clustering, an approach developed in 1985 by Small and Sweeney~\cite{small_clustering_1985}. Variable level clustering involves applying a threshold (below which all edges in a graph are deleted) then iteratively selecting edges with the highest normalized co-citation value and extracting connected components from the graph as clusters for each edge in turn. Three parameters are needed: (i) a threshold or starting level based on a quantile of normalized co-citation frequency, (ii) a level increment, and (iii) a maximum cluster size. An issue is the generation of very large clusters by chaining via low edge weights.  Thus, at each iteration, any cluster exceeding the maximum cluster size  is returned to the process and a higher threshold is applied to break such clusters.

In our implementation of variable level clustering, Figure~\ref{fig:vlc_process}, we first calculated the number of citations accumulated across Scopus for all 2,685,356 articles in \emph{comp}. After discarding those publications without any citations, we restricted further analysis to highly cited articles, those in the 90th percentile of citations or higher, resulting in a dataset of size 212,311. We then identified 4,318,305 publications in Scopus that cite these 212,311 articles from \emph{comp}. For each of the 4,318,305 citing publications in turn, all possible ${n \choose 2}$ reference pairs were generated from a publication's cited references, where $n$ is the number of references in a publication. The cited reference pairs this generated, were then restricted to those in the set of 212,311 highly cited papers previously identified. The frequency of these co-cited pairs was then computed across the \emph{comp} dataset  and normalized using Salton's cosine formula~\cite{salton_citation_1979}. A total of 46,463,117 unique co-cited pairs were thus obtained. These data were represented in a graph where each node was a publication and the weighted edge between the pair was the normalized co-citation frequency. \par

We applied variable clustering with initial parameters of threshold ($t$) = median normalized co-citation frequency (quantile=0.5), increment $i = 0.1$, and maximum cluster size, $mcs=200$. Thus, at start, all edges below the median normalized co-citation frequency were deleted. Clusters were formed by assembling connected components from each co-cited pair beginning with the heaviest edge weight. Clusters below size 100 were retained and any cluster larger than 200 nodes was carried over to the next round. The threshold, $t$, was then incremented by 0.1 and the process repeated while progressively incrementing $t$.  We used a bi-phasic approach where in which $t$ ranged from 0.5-0.9, after which $i$ was reduced to 0.01 for the range $0.9 \leq t \leq 0.99$. A final threshold of $t$=0.999 was applied to break the single remaining large cluster.  Using this approach, 22,232 clusters containing 84,591 nodes were generated, each containing less than 100 nodes. Clusters containing only 2 nodes were discarded bringing the total number of clusters down to 10,298. The publications in these 10,298 clusters were overwhelmingly drawn from the Physical Sciences (one of the 4 top level categories in the Scopus ASJC classification), of which computer science is a sub-category, Figure~\ref{fig:cocitr_venn2}.

@@

In our study of the structure of a twenty year cohort of the computer science literature, we chose to use both traditional journal-based and article-based approaches. Considering that traditional disciplines, `may only partly reflect the actual organization of today’s scientific research'~\cite{waltman_new_2012}  we constructed article clusters at high levels of aggregation using citations in two different ways to examine the \emph{comp} dataset and mapped these clusters to journal-based categories.

For the purpose of this study, our working definition of the computer science literature was all publications in the DBLP bibliography that (i) had a digital object identifier (DOI), and (ii) could also be matched to article identifiers in the Scopus bibliography. Our underlying assumption is that records in DBLP are greatly enriched for computer science even while recognizing that coverage may be incomplete and possibly biased. Cross-matching DBLP publications to records in the Scopus, abstract and citation database of peer-reviewed literature, enabled us to harvest the richer links in Scopus as well as extract links to publications from other disciplines. This cross-matching also allows the use of both journal and article based classifications when clustering documents. Using both direct citations and co-citations as the basis for clustering, we used DBLP articles from journals and conference proceedings for the period 1996-2015. We reconciled these clusters with the All Science Journal Classification (ASJC) developed and maintained by Scopus through a combination of automated and manual procedures. The process of selection and matching resulted in a dataset of 2,685,356 publications (Materials and Methods).

Of these roughly 2.8 million publications, approximately 2.07 million were classified as COMP under subject area and assigned ASJC codes corresponding to Computer Science (major subject area with 13 minor subject areas (Table~\ref{tab:comp}), with the balance of 610,000 publications non-exclusively shared between 26 different major subject areas ranging from Engineering (330,048) to Dentistry (Figure~\ref{fig:ar_cp_annotation}).

The set of 2.07 million publications classified as COMP in Scopus spanned all 13 minor subject areas with Software 30.3\% being the largest component and Computer Science (miscellaneous) 0.9\% the smallest (Table~\ref{tab:comp}). The number of minor subject areas assigned to a publication averaged to 1.75 with 53\% of the publications being assigned to one minor subject area and 0.0003\% assigned to 7 different minor subject areas.

Thus, the sets of computer science publications in DBLP and Scopus are partially overlapping rather than DBLP being a perfect subset of Scopus. [the point being that we should justify our use of both journal and article level classifications].

\subsection{Clustering by Direct Citation} Clustering of publications is commonly accomplished through direct citation, bibliographic coupling, and co-citation with direct citation being proposed as as the best approach to concentrate citation links~\cite{klavans_which_2017}. Accordingly, we first used direct citation links as the basis for cluster formation. A set of 20 clusters was generated using Graclus, a spectral clustering technique that has previously been applied to citation data~\cite{graclus_2007,subelj_clustering_2016}. We constructed article level clusters of this computer science dataset at a sufficiently high level of aggregation to avoid cognitive challenge and cross-matched them to disciplinary classification by journal, the Scopus ASJC classification. To focus on relatively high signal, we only considered Scopus ASJC minor subject area categories that accounted for at least 15\% of the publications in each cluster.

Figure~\ref{fig:heatmap} permits examination of these data from two perspectives: (i) rows: the clusters that map to a given ASJC minor subject area (ii) columns: ASJC minor subject areas that comprised at least 15\% of the publication in a cluster. Under these conditions of clustering and a threshold of 15\%, 31 of the 334 ASJC minor subject areas are represented. Unsurprisingly, the broad categories Computer Science Applications, Software, and Electrical and Electronic Engineering register in 16, 12, and 10 clusters respectively while Artificial Intelligence mapped to 7 different clusters. At the other end of the range, 12 of the 31 ASJC minor subject areas were each detected in a single cluster.

From the alternate perspective (columns), cluster 17 was the most diverse and contained publications annotated with 8 minor subject area labels: Biochemistry, Chemistry(all), Genetics, Molecular Biology,  Statistics \& Probability, Computational Theory \& Mathematics, and Computational Mathematics. Cluster 19 mapped to two areas but was excluded from qualitative analysis because of its high conductance value. Of the remaining clusters, the third (Cluster 2) represents interactions between the minor subject areas Theoretical Computer Science, Discrete Mathematics and Combinatorics, and Applied Mathematics as well as the more generic Computer Science (all). Clusters 3 and 4 include Computer Networks and Communications and Cluster 18 (Artificial intelligence, Cognitive neuroscience, and Neurology) and clusters 5-9 include Management Science, Operations Research, Information Systems, Modeling and Simulation, and Human Computer Interaction.

To us, these data suggest that areas central to computer science in 2019 (Salton and Bergmark's historical triad of hardware, software, and theory) are more likely to be found in multiple clusters, whereas areas such as biology, chemistry, and statistics are more restricted in distribution.  A second inference is that, in some cases journal-based classification and our article clusters are similar in terms of article content (Hardware and Architecture). Thirdly, that the ASJC minor subject area, computer applications is relatively non-specific, and publications thus labeled are near ubiquitous (16/20).  Fourthly, that for the DBLP dataset, interactions with fields outside computer science are restricted to a single cluster or small numbers of clusters, e.g., cluster 17  and cluster 18.

We also explored mapping these 20 clusters assembled through direct citation to the CCS classification by manually matching the top 25-50 most heavily cited publications in each cluster to corresponding categories in the CCS. This was feasible in some cases, clusters 0 and 1 mapped reasonably well to the top level categories Hardware and Computer Systems Organization but much less so in the cases of other clusters. Further, the top cited papers often derived from biology and we were concerned that they were not representative of the cluster. The challenges inherent in labeling citation clusters at a high level of aggregation have been previously noted~\cite{waltman_new_2012}.

[closing sentence necessary?]

\subsection{Clustering by Co-citation.} For an alternate examination of the data, we used co-citation frequencies to cluster the DBLP dataset. Unlike clustering by direct citation with Graclus, where every input publication is assigned to a cluster and every citation is weighted equally, clustering by co-citation is progressively selective, involves assigning weights to co-citations, and also accounts for weak inter-cluster interactions.

As described in Methods, we first identified the top 10\% most-cited articles in the \emph{comp} dataset and then documented all co-cited pairs of their cited references that were also elements of \emph{comp}. The frequency of co-citation was then computed across all records in Scopus and normalized using Salton's correction. These co-cited pairs and the normalized co-citation frequencies were then represented as a graph where each node was a publication and each edge represented a co-citation, weighted by the normalized co-citation frequency.

For clustering, we applied variable level clustering (VLC) followed by 600 iterations of agglomeration of the clusters resultant from VLC since further rounds of agglomeration resulted in the development of a single large cluster (Fig~\ref{fig:merge_select}). To focus on major clusters, we also restricted agglomeration to clusters containing at least 10 nodes.  Consequently, the 20 largest clusters were used for further analysis and comparison with clusters generated by direct citation. The largest of these 20 clusters contained 1,276 nodes and the smallest contained 136, with total nodes summing to 8,042.

Fig~\ref{fig:heatmap_cocit} shows a heatmap in which clusters constructed by co-citation are mapped to Scopus ASJC minor subject areas. Mappings are only shown if they account for at least 15\% of the publications in a cluster. Unlike the earlier figure [insert figure reference] Unlike the case of clustering by direct citation, no cluster maps to more than three ASJC minor subject areas. We have to decide what to say about this figure but I think it suggests intra-cluster relationships for the most part.

>>>>

Computer science and its applications have experienced rapid growth and diversification over the last twenty years. As observed in a 2017 US National Academies Report, ``A wide range of jobs in virtually all sectors demand computing skills to an unprecedented extent. And every academic discipline finds itself incorporating computing into its research and educational mission"~\cite{nas_2017}. More recently, the collective influence of the Internet of Things (IoT), `big' data, accessible cloud computing, and advances in artificial intelligence have been postulated as a recent driver for growth and evolution~\cite{siebel2019_digital}. Given this powerful influence some understanding of the present state and structure of computer science and its relationship to other fields could inform planning and policy making at multiple levels from national level funding all the way down to faculty hiring strategy. For the same reason of broad influence, computer science can also serve as a model system for studying interdisciplinarity.

Noting that that the scientific literature serves a rich source of information to study the structure and historical development of a field, 
Salton and Bergmark conducted a historical study  in 1979 of the computer science literature (419 computer articles published in 1974,  and 3,812 references cited in these articles)~\cite{salton_citation_1979}. These authors described the global structure of computer science as comprising three main areas (i) theoretical foundations, such as theory of computation, (ii) hardware and computer systems, such as architecture,  and (iii) software, such as programming systems.  Related areas noted were (a) mathematics of computing, such as numerical analysis, (b) special software topics, such as operating systems, (c) data management and database systems, (d) methodologies valid for multiple applications, such as algebraic manipulation, (e) computer applications, such as computer graphics, and  (f) non-technical aspects, such as computer education.

Looking beyond Salton and Bergmark's historical triad of theoretical foundations; hardware and computer systems; and software, the Computing Classification System (CCS) published by the Association for Computing Machinery now consists of 13 top-level areas that reflect a more current view of the field~\cite{acm_ref}. This classification also addresses relationships with other fields under the category of Applied Computing. However, an easy way to map scientific publications to the CCS, especially interdisciplinary articles or those from proximal fields, does not seem to be available.

Other classification systems are available such as the All Science Journal Classification (ASJC) developed and maintained by Scopus, the Web of Science research and categorical classifications from Clarivate Analytics, and the NSF classification system~\cite{nsf_classification} [add citations for Scopus and Web of Science]. Scopus and the Web of Science also include comprehensive bibliographies with citation links and proprietary unique identifiers for publications. 
All three systems rely on applying one or more journal-derived labels to articles.  
A logical prediction, given the diversity of articles within journals, is limited specificity at the article level. 
Shu and colleagues recently noted in a comparative study of the Chinese Science Citation Database (CSCD) and the Web of Science that 46\% of articles did not belong to the discipline of the journal they were published in~\cite{shu_comparing_2019}. 
Others have also discussed and critiqued disciplinary assignments using journal-based classifications~\cite{wang_large-scale_2016,perianes-rodriguez_comparison_2017}. 
Article classification systems have been constructed that escape some of the criticisms of journal-based classification~\cite{traag_louvain_2019,boyack_classification_2014,waltman_new_2012}, but seem to be in the research domain rather than the domain of practical use [need to check on this].[May need another bridging sentence to the next paragraph].

@@

 \subsection{Overview}
 In our study of the structure of a twenty year cohort of the computer science literature, we chose to use both traditional journal-based and article-based approaches. Considering that traditional disciplines `may only partly reflect the actual organization of today's scientific research'~\cite{waltman_new_2012}, we constructed article clusters at high levels of aggregation using citations to examine the \emph{comp} dataset and mapped these clusters to journal-based categories.
 Clustering of publications is commonly accomplished through direct citation, bibliographic coupling, and co-citation with direct citation being proposed as as the best approach to concentrate citation links~\cite{klavans_which_2017}. 
Accordingly, we used direct citation links as the basis for cluster formation, and also co-citation to obtain an alternative view.

For the purpose of this study, our working definition of the computer science literature was all publications in the DBLP bibliography that (i) had a digital object identifier (DOI), and (ii) could also be matched to article identifiers in the Scopus bibliography. 
Our underlying assumption is that records in DBLP are greatly enriched for computer science even while recognizing that coverage may be incomplete and possibly biased. Cross-matching DBLP publications to records in the Scopus, abstract and citation database of peer-reviewed literature, enabled us to harvest the richer links in Scopus as well as extract links to publications from other disciplines. This cross-matching also allows the use of both journal and article based classifications when clustering documents. Using both direct citations and co-citations as the basis for clustering, we used DBLP articles from journals and conference proceedings for the period 1996-2015. We reconciled these clusters with the All Science Journal Classification (ASJC) developed and maintained by Scopus through a combination of automated and manual procedures, producing a dataset of  2,685,356 publications.

@@

\subsection{Clustering methods} In applying both clustering by direct citation and by co-citation, we attempted to consider, wherever possible, the criteria articulated by \v{S}ubelj, van Eck, and Waltman~\cite{subelj_clustering_2016}, of (i) the largest cluster should be no more than 10 times the smallest one, (ii) small clusters should be eliminated, (iii) small changes and replicates should yield similar results (``stability"),  (iv) computing time should be minimized where possible, and (v) the clustering should seem reasonable on a qualitative level (``intuitive sensibility").

\paragraph{Clustering by Direct Citation.} Graclus~\cite{graclus_2007} is a spectral graph clustering package that optimizes  various clustering criteria, including normalized cut, ratio cut, and ratio association, that has previously been applied to citation data~\cite{subelj_clustering_2016}. 
We used v1.2 in our experiments. 
The \emph{comp} dataset was formatted as an undirected graph, stored in a file with a header line indicating the number of nodes and edges, and used as input to Graclus, which requires the number of clusters to be formed as an input parameter. In preliminary experiments, we varied the number of clusters to be formed between 10 and 50 clusters (data not shown).  At around 20 clusters, clusters size was relatively stable with the largest cluster containing roughly 10 times the number of nodes in the smallest one~\cite{subelj_clustering_2016}. Consequently, we used Graclus to generate 20 clusters, labeled 0-19 (Table~\ref{tab:graclus}).

We also used conductance, as defined in Shun et al.~\cite{shun_parallel_2016}, to  evaluate clustering (smaller is better), noting that conductance has been found to be ``the stand-alone quality metric that best indicates performance on information recovery metrics" \cite{emmons2016analysis}. 
%Tandy, George - add reference to Zaki paper as well
In our analysis, we saw that the last cluster (cluster 19) had a much larger conductance value than the other clusters and also had the smallest
number of nodes.
We then examined
results obtained using Graclus with two other numbers of clusters (18 and 22), and
in each case, the highest numbered cluster  had the greatest conductance value and also the smallest number of nodes.
These results suggest that Graclus produces a final cluster that effectively serves as a container for `left over publications' during the clustering procedure. 
Therefore, we limited our consideration of cluster 19 (the final of the twenty clusters)  when interpreting results.  The remaining 19 clusters had conductance values ranging from 0.09 to 0.25 with a median conductance of 0.15 (Fig.~\ref{fig:graclus_comparison_conductance}, Table~\ref{tab:comp}).

\paragraph{Clustering by Co-Citation.}
\textcolor{blue}{Check the next paragraphs for repetitiveness.}

For an alternate view of these DBLP data we constructed clusters using co-citation, the frequency with which a pair of articles is cited by other articles. 
Unlike clustering by direct citation with Graclus, where every input publication is assigned to a cluster and every citation is weighted equally, clustering by co-citation is progressively selective, involves assigning weights to co-citations, and also accounts for weak inter-cluster interactions.

We first identified the top 10\% most-cited articles in the \emph{comp} dataset and then documented all co-cited pairs of their cited references that were also elements of \emph{comp}. The frequency of co-citation was then computed across all records in Scopus and normalized using Salton's correction. These co-cited pairs and the normalized co-citation frequencies were then represented as a graph where each node was a publication and each edge represented a co-citation, weighted by the normalized co-citation frequency.

%For clustering, we applied a variant of variable level clustering (VLC, described below) followed by 600 iterations of agglomeration of the clusters resultant from VLC since further rounds of agglomeration resulted in the development of a single large cluster (Fig.~\ref{fig:merge_select}). To focus on major clusters, we also restricted agglomeration to clusters containing at least 10 nodes.  Consequently, the 20 largest clusters were used for further analysis and comparison with clusters generated by direct citation. The largest of these 20 clusters contained 1,276 nodes and the smallest contained 136, with total nodes summing to 8,042.

We used a modification of variable level clustering combined with agglomerative clustering, an approach developed in 1985 by Small and Sweeney~\cite{small_clustering_1985}. Variable level clustering involves applying a threshold (below which all edges in a graph are deleted) then iteratively selecting edges with the highest normalized co-citation value and extracting connected components from the graph as clusters for each edge in turn. Three parameters are needed: (i) a threshold or starting level based on a quantile of normalized co-citation frequency, (ii) a level increment, and (iii) a maximum cluster size. An issue is the generation of very large clusters by chaining via low edge weights.  Thus, at each iteration, any cluster exceeding the maximum cluster size  is returned to the process and a higher threshold is applied to break such clusters.

In our implementation of variable level clustering (Fig.~\ref{fig:vlc_process}), we first calculated the number of citations accumulated across Scopus for all 2,685,356 articles in \emph{comp}. After discarding those publications without any citations, we restricted further analysis to highly cited articles--those in the 90th percentile of citations or higher--resulting in a dataset of size 212,311. We then identified 4,318,305 publications in Scopus that cite these 212,311 articles from \emph{comp}. For each of the 4,318,305 citing publications in turn, all possible ${n \choose 2}$ reference pairs were generated from a publication's cited references, where $n$ is the number of references in a publication. The cited reference pairs this generated were then restricted to those in the set of 212,311 highly cited papers previously identified. The frequency of these co-cited pairs was then computed across the \emph{comp} dataset  and normalized using Salton's cosine formula~\cite{salton_citation_1979}. A total of 46,463,117 unique co-cited pairs were thus obtained. These data were represented in a graph where each node was a publication and the weighted edge between the pair was the normalized co-citation frequency. \par

We applied variable clustering with initial parameters of threshold ($t$) = median normalized co-citation frequency (quantile=0.5), increment $i = 0.1$, and maximum cluster size, $mcs=200$. Thus, at start, all edges below the median normalized co-citation frequency were deleted. Clusters were formed by assembling connected components from each co-cited pair beginning with the heaviest edge weight. Clusters below size 100 were retained and any cluster larger than 200 nodes was carried over to the next round. The threshold, $t$, was then incremented by 0.1 and the process repeated while progressively incrementing $t$.  We used a bi-phasic approach where in which $t$ ranged from 0.5-0.9, after which $i$ was reduced to 0.01 for the range $0.9 \leq t \leq 0.99$. A final threshold of $t$=0.999 was applied to break the single remaining large cluster.  Using this approach, 22,232 clusters containing 84,591 nodes were generated, each containing less than 100 nodes. Clusters containing only 2 nodes were discarded bringing the total number of clusters down to 10,298. The publications in these 10,298 clusters were overwhelmingly drawn from the Physical Sciences (one of the 4 top level categories in the Scopus ASJC classification), of which computer science is a sub-category (Fig.~\ref{fig:cocitr_venn2}).

@@

\subsection{Data}
The process of selection and matching resulted in a dataset of 2,685,356 publications (Materials and Methods). 
Of these roughly 2.8 million publications, approximately 2.07 million were classified as COMP under subject area and assigned ASJC codes corresponding to Computer Science (major subject area with 13 minor subject areas (Table~\ref{tab:comp}), with the balance of 610,000 publications non-exclusively shared between 26 different major subject areas ranging from Engineering (330,048) to Dentistry (Fig.~\ref{fig:ar_cp_annotation}).

The set of 2.07 million publications classified as COMP in Scopus spanned all 13 minor subject areas with Software 30.3\% being the largest component and Computer Science (miscellaneous) 0.9\% the smallest (Table~\ref{tab:comp}). The number of minor subject areas assigned to a publication averaged to 1.75 with 53\% of the publications being assigned to one minor subject area and 0.0003\% assigned to 7 different minor subject areas. 
We note also that  the $4.8x10^6$ DBLP publications are only partially found in Scopus (which had a total of $5.8x10^6$ publications), using our methodology.
%George, please consider revising this text
%the sets of computer science publications in DBLP and Scopus are partially overlapping rather than DBLP being a perfect subset of Scopus. [the point being that we should justify our use of both journal and article level classifications].

\subsection{Clustering by Direct Citation}

We constructed article-level clusters of this computer science dataset at a sufficiently high level of aggregation to avoid cognitive challenge and cross-matched them to  the  Scopus ASJC classification. 
To focus on relatively high signal, we only considered Scopus ASJC minor subject area categories that accounted for at least 15\% of the publications in each cluster.

Figure~\ref{fig:heatmap} permits examination of these data from two perspectives: (i) rows: the clusters that map to a given ASJC minor subject area (ii) columns: ASJC minor subject areas that comprised at least 15\% of the publication in a cluster. Under these conditions of clustering and a threshold of 15\%, 31 of the 334 ASJC minor subject areas are represented. Unsurprisingly, the broad categories Computer Science Applications, Software, and Electrical and Electronic Engineering register in 16, 12, and 10 clusters respectively, while Artificial Intelligence mapped to 7 different clusters. 
At the other end of the range, 12 of the 31 ASJC minor subject areas were each detected in a single cluster.

From the alternate perspective (columns), cluster 17 was the most diverse and contained publications annotated with 8 minor subject area labels: Biochemistry, Chemistry(all), Genetics, Molecular Biology,  Statistics \& Probability, Computational Theory \& Mathematics, and Computational Mathematics. Cluster 19 mapped to two areas but was excluded from qualitative analysis because of its high conductance value. Of the remaining clusters, the third (Cluster 2) represents interactions between the minor subject areas Theoretical Computer Science, Discrete Mathematics and Combinatorics, and Applied Mathematics as well as the more generic Computer Science (all). Clusters 3 and 4 include Computer Networks and Communications and Cluster 18 (Artificial Intelligence, Cognitive Neuroscience, and Neurology) and clusters 5--9 include Management Science, Operations Research, Information Systems, Modeling and Simulation, and Human-Computer Interaction.

These data suggest that areas central to computer science in 2019 (Salton and Bergmark's historical triad of hardware, software, and theory) are more likely to be found in multiple clusters than peripheral topics.
%, whereas areas such as biology, chemistry, and statistics are more restricted in distribution.  
A second inference is that, in some cases, journal-based classification and our article clusters are similar in terms of article content (Hardware and Architecture). 
A third inference is that the ASJC minor subject area ``Computer Applications" is relatively non-specific, and publications thus labeled are nearly ubiquitous (16/20).  Finally,   for the DBLP dataset, interactions with fields outside computer science (e.g., Biochemistry)  are  typically restricted to a single cluster or found in a small number of clusters.

We also note that these clusters cannot be adequately characterized by using a single CCS classification, a challenge that has been previously noted \cite{waltman_new_2012}.
For example, we
manually matched the top 25--50 most heavily cited publications in each cluster to corresponding categories in the CCS. This was feasible in some cases, but not in all. For example, clusters 0 and 1 mapped reasonably well to the top level categories Hardware and Computer Systems Organization,  but 
the top cited papers often derived from biology and yet biology was clearly not  representative of all of these affected clusters.

\subsection{Clustering by Co-citation.} For an alternate examination of the data, we used co-citation frequencies to cluster the DBLP dataset. 
Figure \ref{fig:heatmap_cocit} shows a heatmap in which clusters constructed by co-citation are mapped to Scopus ASJC minor subject areas. Mappings are only shown if they account for at least 15\% of the publications in a cluster. 
\textcolor{blue}{
Unlike the earlier figure [insert figure reference] Unlike the case of clustering by direct citation, no cluster maps to more than three ASJC minor subject areas. We have to decide what to say about this figure but I think it suggests intra-cluster relationships for the most part. }
